Predicting Which States are Most Interested in the term *China Virus* Based off Identity Politics, Demography and COVID-19 Statistics 

### 2020-05-05 Final-Draft.Rmd
finishing intro, addding table, cleaning doc

On March 17, 2020 President Trump referred to the Coronavirus as the "China Virus." Shortly after, there was an uptick in verbal and physical attacks against Asian-Americans. One aspect of public health that is often thrown to the wayside is how influential public officials and leaders are in disseminating public health information. Moreover, not only can their words change the public's views on a health matter but it can also shift a nation's perspective on someone's identity. The Asian Pacific Policy and Planning Council reported that in their first four weeks of receiving reports on COVID-19 anti-Asian discirimnation, they recieved more than 1,500 incidents [CITE]. This abrupt spasm of racism mimics the kind faced by American Muslims, Arabs and South Asians in the United States after the terrorist attacks of 9/11. However, when President George W. Bush urged tolerance of American Muslims, this time President Trump is using language that Asian-Americans say is inciting racist attacks.

In addition, given the influence of identity politics we may expect the term "China Virus" to be more polarizing to certain identities and states.The Pew Reserach Center discovered that nearly three-quarters of Repbulicans and Republican-leaning indepdents view China unfavoriably[CITE]. But the majority of reports to the AAPI were from Democrat leaning states like California, Washington and Illinois [CITE]. Moreover, states that have a higher number of reported positive cases tend to have a higher interest in Googling the term "China Virus" [CITE].  Thus, this begs the question: What factors of the poltiical, demographic and COVID-19 spheres best predicts the level of interest a state will have in the term "China Virus?"

In order to answer this question, we utilized Monte Carlo Markov Chains to build four predictive models that included some of the variables mentioned above. Model 1 is our simplest model while Model 4 is our most complex model. We then ran several different diagnostic tests such as MCMC and PP checks to evaluate the quality and structures of our models. While all the models had some different levels of predictive strength, our best model was ___________. This model proved to provide the best balance between complexity and quality.

\

@@

## 2.2 Variables of interest

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}
b<- ggplot(Finaldata, aes(x = ChinaVirusInterest, fill = as.factor(State))) + 
  geom_density(alpha = 0.5)+ ggtitle("China Virus Density by State") + labs(fill="State")

ggplotly(b)
```
As we can see from the density plot of the China Virus Interest during our time period 2020-03-14 - 2020-03-21, it behaves relatively normal with a small bump at 0. As a group we believe this bump occurs as 0 is the lowest value it can take and because of that limitation of the China Virus Interest we see a small bump around 0. One could argue against a normal distribution as it kinda looks a bit right skewed. But our team believes that a normal distribution is the best at describing the density.

@@

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}
c <-  ggplot(Finaldata, aes(y=positive, x=Winner, fill=(Winner)))+
      geom_boxplot() +
      facet_wrap(~Region)+
      theme(legend.position = "none") +
      scale_fill_manual(values = c("Democrat"="blue3", "Republican"= "red3"))+
      lims(y= c(0,1000))+
      ggtitle("Positive COVID-19 Cases by State Political Party Winner & Region")
ggplotly(c)
```
This visualisation depicts the distribution of positive COVID-19 cases by region and by which political party won in the 2016 elections. We can see that Democrat states in the Midwest, Mountain, and West have a larger range and higher quantile metrics for positive cases overall. For the Northeast and South regions the mean of positive COVID-19 cases are higher but not significantly. This is an interesting pattern considering that poltical party affiliation appears to interact with the number of postive cases by region.

@@

## Model 1 Repeated Measures Model

@@

### Model Structure

@@

## Model2: Normal Regression

@@

### Model Structure
$$\begin{aligned}
Y_{i}|\beta_0, \beta_1, \beta_2,\beta_3,\beta_4 &\overset{ind}{\sim} N(\beta_0+ \beta_1X_1 + \beta_2X_2 + \beta_3X_3+ \beta_4X_4,\sigma^2)\\
\beta_0,\beta_1, \beta_2,\beta_3 &\sim N(...,...)\\
\sigma &\sim Exp(...)\\
\end{aligned}$$

@@

## Model 3: Normal Regression + Repeated Measures

For our third model, we decided to combine the first two models to first understand variations in political, demographic, covid-related variables and also be able to capture variation that we could not explain in model 2. Additionally, we kept the correlation structure that we saw in model 1, but we included the variables in model 2 so that we could explain why some states had a larger `ChinaVirusInterest` than others. Similar to model 1, model 3 allows us to capture differences in `ChinaVirusInterest` with the $\theta_i$ value which represents each state's mean value.
### Model Structure

@@

## Model 4: Longitudinal Model (state-specific slopes)

@@

### Model Structure

@@

##mcmc plots
```{r}
mcmc_trace(RM_model, pars = c("(Intercept)","Sigma[State:(Intercept),(Intercept)]"))
mcmc_dens_overlay(RM_model, pars = c("(Intercept)","Sigma[State:(Intercept),(Intercept)]")) 
```

@@

##ppcheck

```{r}
pp_check(RM_model, nreps=50)
```

@@

##mcmc plots

@@

##ppcheck

```{r}
pp_check(NR_model, nreps = 50)
```

@@

```{r}
mcmc_trace(model_3, pars=c("(Intercept)","percent_white","StateColorred","StateColorpurple","positive","Day","b[(Intercept) State:AK]","b[(Intercept) State:WA]"),facet_args = list(ncol = 3, strip.position = "left"))

@@

##ppcheck
```{r}
pp_check(model_3, nreps=50)
```
As seen in our prior posterior predictive checks, they have similar errors around the tails. But it seems that our GLM structure is also reasonable for this model. 
## 4.4 Model 4 Evaluation (Longitudinal)

##mcmc plots

```{r}
# Trace plots
mcmc_trace(complexmod,pars = c("(Intercept)","percent_white","StateColorred","StateColorpurple","positive","Day","b[Day State:AK]","b[Day State:WA]"),facet_args = list(ncol = 3, strip.position = "left")) 
mcmc_trace(complexmod, pars = c("sigma"),regex_pars = "Sigma",facet_args = list(ncol = 2, strip.position = "left"))

@@

##ppcheck

```{r}
pp_check(complexmod, nreps = 50)
```
Overall all ppchecks tend to have the similar graph because we are assuming a normal relationship on the response variable for all of them.

## Are our models strong?: Prediction & classification summaries

@@

```{r, message=FALSE, include=FALSE}
#Repeated Measures
set.seed(454)
pred_1 <- posterior_predict(
  RM_model,
  newdata = model_data, transform = TRUE)

@@

```{r, message=FALSE, include=FALSE}
#Normal Regression
set.seed(454)
pred_2 <- posterior_predict(
  NR_model,
  newdata = model_data, transform = TRUE)

@@

```{r, message=FALSE, include=FALSE}
#Repeated Measures + Normal Regression
set.seed(454)
pred_3 <- posterior_predict(
  model_3,
  newdata = model_data, transform = TRUE)

@@

```{r message=FALSE, include=FALSE}
#Longitudinal Model
set.seed(454)
pred_4 <- posterior_predict(
  complexmod,
  newdata = model_data, transform = TRUE)

@@

### 5.1.a Table

```{r}
all_summ<-rbind(sum1,sum2,sum3,sum4)
rownames(all_summ)<- c("Model 1","Model 2","Model 3","Model 4")
all_summ
```

@@

### Why we chose Model 3 as our best model?

@@

## 5.2 Posterior Prediction One State

```{r}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive_rate = 0.11) #Doesn't recognize state
set.seed(454)
pred_FL_RM <- posterior_predict(
  RM_model,
  newdata = new_FL)

@@

```{r}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive = 658, StateColor="purple") 
set.seed(454)
pred_FL_NR <- posterior_predict(
  NR_model,
  newdata = new_FL)

@@

```{r}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive = 658, StateColor="purple")
set.seed(454)
pred_FL_3 <- posterior_predict(
  model_3,
  newdata = new_FL)

@@

```{r}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive = 658, StateColor="purple")
set.seed(454)
pred_FL_complex <- posterior_predict(
  complexmod,
  newdata = new_FL)

@@

```{r}
actual_FL_pred<-Finaldata %>% select(c(State, ChinaVirusInterest)) %>% filter(State =="FL") %>% summarise(mean=mean(ChinaVirusInterest))
actual_FL_pred
```

@@

# 7. Acknowledgments and References

>>>>

## 1.1 Background

On March 17, 2020 President Trump referred to the Coronavirus as the "China Virus." Shortly after, throughout the subsequent weeks, there was an uptick in verbal and physical attacks against Asian-Americans (Olson, 2020). One aspect of public health that is often thrown to the wayside is how influential public officials and leaders are in disseminating public health information. Moreover, not only can their words change the public's views on a health matter but it can also shift a nation's perspective on someone's identity. The Asian Pacific Policy and Planning Council reported that in their first four weeks of receiving reports on COVID-19 anti-Asian discirimnation, they recieved more than 1,500 incidents (Jeung & Nham 2020). This abrupt spasm of racism mimics the kind faced by American Muslims, Arabs and South Asians in the United States after the terrorist attacks of 9/11 (Tavernise & Oppel Jr, 2020). However, when President George W. Bush urged tolerance of American Muslims, this time President Trump is using language that Asian-Americans say is inciting racist attacks.

In addition, given the influence of identity politics we may expect the term "China Virus" to be more polarizing to certain identities and states.The Pew Reserach Center discovered that nearly three-quarters of Repbulicans and Republican-leaning indepdents view China unfavoriably (Devlin, et al 2020. But the majority of reports to the AAPI were from Democrat leaning states like California, Washington and Illinois [Jeung & Nham 2020]. Moreover, states that have a higher number of reported positive cases tend to have a higher interest in Googling the term "China Virus" (2020).  Thus, this begs the question: What factors of the poltiical, demographic and COVID-19 spheres best predicts the level of interest a state will have in the term "China Virus?"

## 1.2 Brief Synopsis
In order to answer this question, we utilized Monte Carlo Markov Chains to build four predictive models that included some of the variables mentioned above. Model 1, a repeated measures model, is our simplest model, Model 2 is a normal regression based, Model 3 is a combination of normal and repeated measures, and Model 4, our most complex model, is longitudinal based. We then ran several different diagnostic tests such as MCMC and PP checks to evaluate the quality and structures of our models. While all the models had some different levels of predictive strength, our best model was a normal regression and repeated measures model, Model 3. This model proved to provide the best balance between complexity and quality.Model 3 captures important attributes that the normal model and repeated measures did not and it achieves similar results to our complex model, Model 4, without the heavy computational toll. Overall, we hope this project inspires others to explore the intereconnectdness of health, communication, politics and demography.

@@

## 2.2 Variables of interest
+------------------------+---------------------------------------------------+
| Variables:             | Description:                                      |
+========================+===================================================+
| `State`                | State abbreivation and main identifier            |
+------------------------+---------------------------------------------------+
| `StateColor`           | Political Leaning of the state. Either red, blue  |
|                        | or purple.                                        |
+------------------------+---------------------------------------------------+
| `percent_white`        | Percent of the Population that is White           |
+------------------------+---------------------------------------------------+
| `percent_asian`        | Percent of the Population that is Asian           |  
+------------------------+---------------------------------------------------+
|`Positive`              | Number of reported positive COVID-19 cases        | 
+------------------------+---------------------------------------------------+
|`Negative`              | Number of reported negative COVID-19 cases        | 
+------------------------+---------------------------------------------------+
|`Day`                   | Date of report                                    | 
+------------------------+---------------------------------------------------+
|`totalTestResults`      | Total Number test results (Positive +Negative)    | 
+------------------------+---------------------------------------------------+
|`ChinaVirusInterest`    | Interest index from Google searches by state.     |
|                        | Peak search day=100, all other days in set are    |
|                        | based searches on relative to this peak day.      | 
+------------------------+---------------------------------------------------+
|`Region`                | States divided into five different regions:       |
|                        | West, South, Mountain, Northeast, Midwest         |
+------------------------+---------------------------------------------------+

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE, fig.align= center}

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE, fig.align= center}
b<- ggplot(Finaldata, aes(x = ChinaVirusInterest, fill = as.factor(State))) + 
  geom_density(alpha = 0.5)+ ggtitle("China Virus Density by State") + labs(fill="State")

ggplotly(b)
```

As we can see from the density plot of the China Virus Interest during our time period 2020-03-14 - 2020-03-21, it behaves relatively normal with a small bump at 0. As a group we believe this bump occurs as 0 is the lowest value it can take and because of that limitation of the China Virus Interest we see a small bump around 0. One could argue against a normal distribution as it kinda looks a bit right skewed. But our team believes that a normal distribution is the best at describing the density.

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE, fig.align= center}
IAtoMT <- Finaldata %>% select(c(State,ChinaVirusInterest,Day)) %>% filter (State %in% c("IA","ID","IL","LA","MA","MD","ME","MI","MN","MO","MS","MT"))

plot1<- ggplot(IAtoMT, aes(x = Day, y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    theme(legend.position = "none") +
    facet_wrap(~State)

ggplotly(plot1)
```

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE, fig.align= center}
c <-  ggplot(Finaldata, aes(y=positive, x=Winner, fill=(Winner)))+
      geom_boxplot() +
      facet_wrap(~Region)+
      theme(legend.position = "none") +
      scale_fill_manual(values = c("Democrat"="blue3", "Republican"= "red3"))+
      lims(y= c(0,1000))+
      ggtitle("Positive COVID-19 Cases by State Political Party Winner & Region")
ggplotly(c)
```

This visualisation depicts the distribution of positive COVID-19 cases by region and by which political party won in the 2016 elections. We can see that Democrat states in the Midwest, Mountain, and West have a larger range and higher quantile metrics for positive cases overall. For the Northeast and South regions the mean of positive COVID-19 cases are higher but not significantly. This is an interesting pattern considering that poltical party affiliation appears to interact with the number of postive cases by region.

@@

## 3.1 Model 1 Repeated Measures Model

@@

### 3.1a Model Structure

@@

## 3. 2 Model 2: Normal Regression

@@

### 3.2a Model Structure
$$\begin{aligned}
Y_{i}|\beta_0, \beta_1, \beta_2,\beta_3,\beta_4 &\overset{ind}{\sim} N(\beta_0+ \beta_1X_1 + \beta_2X_2 + \beta_3X_3+ \beta_4X_4,\sigma^2)\\
\beta_0,\beta_1, \beta_2,\beta_3 &\sim N(...,...)\\
\sigma &\sim Exp(...)\\
\end{aligned}$$

@@

## 3.3 Model 3: Normal Regression + Repeated Measures

For our third model, we decided to combine the first two models to first understand variations in political, demographic, covid-related variables and also be able to capture variation that we could not explain in model 2. Additionally, we kept the correlation structure that we saw in model 1, but we included the variables in model 2 so that we could explain why some states had a larger `ChinaVirusInterest` than others. Similar to model 1, model 3 allows us to capture differences in `ChinaVirusInterest` with the $\theta_i$ value which represents each state's mean value.

### 3.3a Model Structure

@@

## 3.4 Model 4: Longitudinal Model (state-specific slopes)

@@

### 3.4a Model Structure

@@

### 4.1a MCMC Plots
```{r, echo=FALSE, warning=FALSE}
mcmc_trace(RM_model, pars = c("(Intercept)","Sigma[State:(Intercept),(Intercept)]"))
mcmc_dens_overlay(RM_model, pars = c("(Intercept)","Sigma[State:(Intercept),(Intercept)]")) 
```

@@

### 4.1b Posterior Predictive Check

```{r, echo=FALSE, warning=FALSE}
pp_check(RM_model, nreps=50)
```

@@

### 4.2a MCMC Plots

@@

### 4.2b Posterior Predictive Check

```{r, echo=FALSE, warning=FALSE}
pp_check(NR_model, nreps = 50)
```

@@

### 4.3a MCMC Plots
```{r, echo=FALSE, warning=FALSE}
mcmc_trace(model_3, pars=c("(Intercept)","percent_white","StateColorred","StateColorpurple","positive","Day","b[(Intercept) State:AK]","b[(Intercept) State:WA]"),facet_args = list(ncol = 3, strip.position = "left"))

@@

### 4.3b Posterior Predictive Check 
```{r, echo=FALSE, warning=FALSE}
pp_check(model_3, nreps=50)
```
As seen in our prior posterior predictive checks, they have similar errors around the tails. But it seems that our GLM structure is also reasonable for this model.

## 4.4 Model 4 Evaluation (Longitudinal)

### 4.4a MCMC Plots

```{r, echo=FALSE, warning=FALSE}
# Trace plots
mcmc_trace(complexmod,pars = c("(Intercept)","percent_white","StateColorred","StateColorpurple","positive","Day","b[Day State:AK]","b[Day State:WA]"),facet_args = list(ncol = 3, strip.position = "left")) 
mcmc_trace(complexmod, pars = c("sigma"),regex_pars = "Sigma",facet_args = list(ncol = 2, strip.position = "left"))

@@

### 4.4b Posterior Predictive Check

```{r, echo=FALSE, warning=FALSE}
pp_check(complexmod, nreps = 50)
```
Overall all ppchecks tend to have the similar graph because we are assuming a normal relationship on the response variable for all of them.

## 4.5 Are our models strong?: Prediction & classification summaries

@@

### 5.1a 
```{r, message=FALSE, include=FALSE}
#Repeated Measures
set.seed(454)
pred_1 <- posterior_predict(
  RM_model,
  newdata = model_data, transform = TRUE)

@@

### 5.1b
```{r, message=FALSE, include=FALSE}
#Normal Regression
set.seed(454)
pred_2 <- posterior_predict(
  NR_model,
  newdata = model_data, transform = TRUE)

@@

### 5.1c
```{r, message=FALSE, include=FALSE}
#Repeated Measures + Normal Regression
set.seed(454)
pred_3 <- posterior_predict(
  model_3,
  newdata = model_data, transform = TRUE)

@@

### 5.1d
```{r message=FALSE, include=FALSE}
#Longitudinal Model
set.seed(454)
pred_4 <- posterior_predict(
  complexmod,
  newdata = model_data, transform = TRUE)

@@

### 5.2 Table

```{r, echo=FALSE, warning=FALSE}
all_summ<-rbind(sum1,sum2,sum3,sum4)
rownames(all_summ)<- c("Model 1","Model 2","Model 3","Model 4")
all_summ
```

@@

### 5.2a Model 3 is the best

@@

## 5.3 Posterior Prediction One State

### 5.3a
```{r, echo=FALSE, warning=FALSE}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive_rate = 0.11) #Doesn't recognize state
set.seed(454)
pred_FL_RM <- posterior_predict(
  RM_model,
  newdata = new_FL)

@@

### 5.3b
```{r, echo=FALSE, warning=FALSE}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive = 658, StateColor="purple") 
set.seed(454)
pred_FL_NR <- posterior_predict(
  NR_model,
  newdata = new_FL)

@@

###5.3c
```{r, echo=FALSE, warning=FALSE}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive = 658, StateColor="purple")
set.seed(454)
pred_FL_3 <- posterior_predict(
  model_3,
  newdata = new_FL)

@@

### 5.3d
```{r, echo=FALSE, warning=FALSE}
### For predicting one case 
new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive = 658, StateColor="purple")
set.seed(454)
pred_FL_complex <- posterior_predict(
  complexmod,
  newdata = new_FL)

@@

### 5.3e
```{r, echo=FALSE, warning=FALSE}
actual_FL_pred<-Finaldata %>% select(c(State, ChinaVirusInterest)) %>% filter(State =="FL") %>% summarise(mean=mean(ChinaVirusInterest))
actual_FL_pred
```

@@

# 7. Acknowledgments

# 8. References
Devlin, K., Silver, L., Huang, C. (2020). Amid Coronavirus Outbreak, Americans' Views of China Increasingly Negative. Retrieved 5 May 2020, from https://www.pewresearch.org/global/2020/04/21/u-s-views-of-china-increasingly-negative-amid-coronavirus-outbreak/

Google Trends. (2020). Retrieved 5 May 2020, from https://trends.google.com/trends/explore?geo=US&q=china%20virus

Jeung, R. Nham, K. (2020). STOP AAPI HATE MONTHLY REPORT 4/23/20. Retrieved 5 May 2020, from http://www.asianpacificpolicyandplanningcouncil.org/wp-content/uploads/STOP_AAPI_HATE_MONTHLY_REPORT_4_23_20.pdf

Olson, H. (2020). Trump’s not the only one blaming China. Americans increasingly are, too. Retrieved 5 May 2020, from https://www.washingtonpost.com/opinions/2020/05/04/get-ready-an-election-all-about-china/

Tavernise, S. Oppel Jr, R. (2020). Spit On, Yelled At, Attacked: Chinese-Americans Fear for Their Safety. Retrieved 5 May 2020, from https://www.nytimes.com/2020/03/23/us/chinese-coronavirus-racist-attacks.html
